Supplementary Figures relative to dimensionality reduction

### 2020-01-29 supplementary_figures/supp1_dataset_description.Rmd
updated supplementary figures

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(SingleCellExperiment)
  library(scater)
  library(reshape2)
  library(pheatmap)
})
suppressMessages(devtools::load_all("../../pipComp/"))

# we load the datasets
lf <- list.files("../datasets/", full.names=TRUE)
names(lf) <- gsub(".SCE.rds", "", basename(lf), fixed=TRUE)
SCEs <- lapply(lf, readRDS)
```

@@

```{r misclass, fig.height=4}
theme_set(theme_cowplot(font_size = 11))
mis <- readRDS("../data/misclassification.rds")
pdb <- function(x, ...){
  n <- names(mis)[x]
  plotBadCells(SCEs[[n]], mis[[n]], name=n, ...)
}
pdb(2, labels=c("A",rep("",6)))
```
\vfill
```{r, fig.height=4}
pdb(3, labels=c("B",rep("",6)))
```

>>>>

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(SingleCellExperiment)
  library(scater)
  library(reshape2)
  library(pheatmap)
})
suppressMessages(devtools::load_all("../../pipComp/"))
theme_set(theme_cowplot(font_size = 12))
# we load the datasets
lf <- list.files("../datasets/", full.names=TRUE)
names(lf) <- gsub(".SCE.rds", "", basename(lf), fixed=TRUE)
lf <- lf[grep("simMix", lf, invert=TRUE)]
SCEs <- lapply(lf, readRDS)
```

@@

```{r misclass, fig.height=4}
theme_set(theme_cowplot(font_size = 10))
mis <- readRDS("../data/misclassification.rds")
pdb <- function(x, ...){
  n <- names(mis)[x]
  plotBadCells(SCEs[[n]], mis[[n]], name=n, ...)
}
pdb(2, labels=c("A",rep("",6)))
```
\vfill
```{r, fig.height=4}
pdb(3, labels=c("B",rep("",6)))
```

### 2020-01-29 supplementary_figures/supp2_filtering.Rmd
updated supplementary figures

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(SingleCellExperiment)
  library(reshape2)
  library(ComplexHeatmap)
  library(knitr)
})
suppressMessages(devtools::load_all("../../pipComp/"))
source("../misc_functions.R")
data(datasets)

# we load the datasets
lf <- list.files("../datasets/", full.names=TRUE)
names(lf) <- gsub(".SCE.rds", "", basename(lf), fixed=TRUE)
SCEs <- lapply(lf, readRDS)
```

@@

```{r filt_more_ari, fig.width=8, fig.height=8.5}
res <- readRDS("../data/filtering_old.rds")
res$pcCells <- round(100*res$nbCells/datasets[as.character(res$dataset),"nbCells"],2)
# we restrict to analyses that gave a number of clusters close to reality:
res2 <- res[which(abs(res$nbClusters-datasets[as.character(res$dataset), "subpopulations"])<=1),]
p1 <- ggplot(res2, aes(x=pcCells, y=ARI, colour=norm)) + geom_point() + facet_wrap(~dataset, scale="free")

fres <- readRDS("../../resNew/filtering_intermediateSummary.rds")
pc <- fres$filtering$pcCells
pc <- pc[grep("doublet",row.names(pc),invert=T),]
row.names(pc) <- reduceMethodNames(row.names(pc),2)
ha <- HeatmapAnnotation(dataset=reduceMethodNames(colnames(pc),1))
shortDSNames <- c("mixology10x3cl"="mix3cl", "mixology10x5cl"="mix5cl", "Zhengmix4eq"="Zh4eq", "Zhengmix4uneq"="Zh4uneq", "Zhengmix8eq"="Zh8eq")
colnames(pc) <- paste(shortDSNames[reduceMethodNames(colnames(pc),1)], reduceMethodNames(colnames(pc),2))
p2 <- grid.grabExpr(draw(Heatmap(pc, name="% out", bottom_annotation=ha)))

plot_grid(p1, p2, labels=LETTERS[1:2], nrow=2)
```

### Supplementary Figure `r getFigNb()`
**A:** Adjusted rand index (ARI) of various clustering pipelines using various combinations of filtering criteria. Only clustering analyses with a number of clusters +/-1 from the real number of clusters were considered. In general, filtering more does not result in more accurate clustering. **B:** Percentage of cells filtered out per subpopulation, using various combination of filters (see methods).

@@

```{r, fig.height=4}
res$filt.times <- sapply(strsplit(res$filt,"_"), FUN=function(x){ if(length(x)<3) return(1); as.numeric(x[[3]])})
res$filt.MADs <- sapply(strsplit(res$filt,"_"), FUN=function(x){ as.numeric(x[[2]]) })
filt.vars <- strsplit(sapply(strsplit(res$filt,"_"), FUN=function(x) x[[1]]),";")
fv <- unique(unlist(filt.vars))
for(f in fv[grep("both",fv,invert=TRUE)]){
  f2 <- gsub("lower$|higher$","both",f)
  res[[f]] <- sapply(filt.vars, FUN=function(x){ any(c(f,f2) %in% x) })
}
form <- paste0("~nbClusters*dataset+norm+",paste(grep(".",colnames(res),fixed=TRUE,value=TRUE),collapse="+"))
form <- gsub("filt.MADs","factor(filt.MADs)",form,fixed=T)
form <- gsub("filt.times","factor(filt.times)",form,fixed=T)
form <- paste0("ARI",form)
message(strwrap(paste("lm(", form, ")")))
mod <- lm(as.formula(form), data=res)
mod <- coef(summary(mod))[grep(".",row.names(coef(summary(mod))),fixed=T),]
colnames(mod)[4] <- "p value"
kable(mod)
```

### Supplementary Figure `r getFigNb()`
Linear regression analysis of ARI on various filtering criteria, correcting for the effect of the number of clusters. Increasing the number of distributions on which a cell must deviate in order to be excluded (`filt.times`) tended to be positively associated with ARI, while increasing the number of median absolute deviations required to be an outlier (`filt.MADs`) was associated with a decrease in ARI.

```{r}
sw <- fres$dimreduction$clust.avg.silwidth[[1]]
sw <- sw[,grep("doublet",colnames(sw))]
w <- grep("seuratvst",colnames(sw))
colnames(sw) <- reduceMethodNames(colnames(sw),2)

chm(list("standard Seurat norm"=sw[,-w], "sctransform"=sw[,w]), scale="row", scaleTitle="z-score", row_ann=data.frame(meanSilWidth=rowMeans(sw), dataset=reduceMethodNames(row.names(sw),1)), show_row_names=FALSE)
```

```{r fig.height=3}
r <- fres$dimreduction$PC.R2[which(fres$dimreduction$PC.R2$dim<=10),]
r <- r[grep("doublet",r$method),]
r <- r[grep("clust",r$method, invert=T),]
r <- aggregate(r$value, by=r[,c(1,3)], FUN=mean)
r <- t(cast2(r, dataset~method, value.var="x", fun.aggregate=mean))
w <- grep("seuratvst", row.names(r))
row.names(r) <- reduceMethodNames(row.names(r), 2)
chm(list("standard Seurat norm"=t(r[-w,]), "sctransform"=t(r[w,])), sameScale=TRUE, 
    scale="row", scaleTitle="z-score", value_format="%.2f")
```

### Supplementary Figure `r getFigNb()`
Impact of the set of filters on the average silhouette width of the subpopulations (top) and on the proportion of variance in the first components explained by clusters (bottom).

```{r}
res <- readRDS("../data/selection_old.rds")
res <- res[grep("2000$",res$sel),]
res <- res[grep("deviance|seurat.variance.standardized",res$sel),]
res$filt <- as.factor(res$filt)
levels(res$filt) <- c("coding only", "no mitochondrial", "no ribosomal", "all")
res2 <- res[which(abs(res$nbClusters-datasets[res$dataset,"subpopulations"])<=1),]
ll <- lapply(split(res2, res2$norm), FUN=function(x){
  m2 <- cast2(x, formula=filt~dataset, fun.aggregate=median, value.var="ARI")
})
chm(ll, scale="column", scaleTitle="z-score", sameScale=TRUE, value_format="%.2f", row_sortFn=function(x){ median(x,na.rm=T) }, cluster_columns=FALSE)
```

>>>>

```{r, include=FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(SingleCellExperiment)
  library(reshape2)
  library(ComplexHeatmap)
  library(knitr)
  library(circlize)
})
suppressMessages(devtools::load_all("../../pipeComp/"))
source("../misc_functions.R")
load("../../pipComp/data/datasets.RData")
```

@@

```{r filt_more, fig.width=8, fig.height=9}
res <- readRDS("../data/results_filtering.rds")
m <- scrna_evalPlot_filtering(res, returnTable=TRUE)
m <- m[grep("veryStringent", m$method, invert=TRUE),]

# # horizontal version
# m <- m[grep("veryStringent", m$method, invert=TRUE),]
# m.f1 <- reshape2::dcast(m, method~dataset, value.var="F1atK")
# n <- sapply(strsplit(gsub("=",".",m.f1[,1]),"\\."),FUN=function(x) rev(x)[1])
# n[grep("=doublet", m.f1[,1])] <- paste0("doubletRemoval+",n[grep("=doublet", m.f1[,1])])
# m.pc <- reshape2::dcast(m, method~dataset, value.var="maxPCout")[,-1]
# m.f1m <- reshape2::dcast(m, method~dataset, value.var="meanF1")[,-1]
# m.pcm <- reshape2::dcast(m, method~dataset, value.var="medianPCout")[,-1]
# m.f1 <- m.f1[,-1]
# row.names(m.f1m) <- row.names(m.pcm) <- row.names(m.f1) <- row.names(m.pc) <- n
# 
# chm(list("mean F1"=m.f1m, "F1 at true \n# clusters"=m.f1, "max % lost per\nsubpopulation"=m.pc, "median % lost per\nsubpopulation"=m.pcm), scale = "column", colors = viridis::inferno(3), value_format = c("%.2f","%.2f","%.0f","%.0f"), scaleFn=list(scale,scale, scale2,scale2), cluster_columns = FALSE,  row_sortFn = function(x, na.rm){ sum(x[1:18],na.rm=TRUE)-median(x[19:36], na.rm=TRUE)/100})

mycast <- function(x, val, form=dataset~method){ 
  x <- reshape2::dcast(x, form, value.var=val)
  row.names(x) <- x[,1]
  as.matrix(x[,-1])
}
m.f1 <- mycast(m, "F1atK")
n <- sapply(strsplit(gsub("=",".",colnames(m.f1)),"\\."),FUN=function(x) rev(x)[1])
n[grep("=doublet", colnames(m.f1))] <- paste0("doubletRemoval+",n[grep("=doublet", colnames(m.f1))])
m.pc <- mycast(m, "maxPCout")
colnames(m.f1) <- colnames(m.pc) <- n
o <- order(colMeans(t(scale(t(m.f1))), na.rm=TRUE)*100-colMeans(t(scale(t(m.pc))), na.rm=TRUE))
m.f1 <- m.f1[,o]
m.pc <- m.pc[,o]
m.f1m <- mycast(m, "meanF1")[,o]
m.pcm <- mycast(m, "medianPCout")[,o]
colnames(m.f1m) <- colnames(m.pcm) <- colnames(m.f1)

names(nn) <- nn <- c("m.f1m","m.f1","m.pc","m.pcm")
cellfns <- lapply(nn, FUN=function(mname){
  if(grepl("f1",mname)){
    sp <- paste0('sprintf("%.1f", 100*',mname,'[i, j])')
  }else{
    sp <- paste0('sprintf("%.1f", ',mname,'[i, j])')
  }
  eval(parse(text=paste0('function(j, i, x, y, width, height, fill){',
      'lab <- ', sp, '; grid.text(lab, x, y, gp = gpar(fontsize = 10)) }')))
})

names(nn) <- c("mean F1", "F1 at true \n# clusters", "max % lost per\nsubpopulation", "median % lost per\nsubpopulation")
hs <- lapply(seq_along(nn), FUN=function(i){
  mat <- get(nn[[i]])
  if(grepl("f1",nn[[i]])){
    mat <- t(scale(t(mat)))
  }else{
    mat <- t(scale(t(sqrt(mat))))
  }
  Heatmap(mat, row_title=names(nn)[[i]], col=viridis::inferno(100), cell_fun = cellfns[[i]], show_heatmap_legend = FALSE, cluster_columns = FALSE, cluster_rows = FALSE)
})

hs[[1]] %v% hs[[2]] %v% hs[[3]] %v% hs[[4]]
```

### Supplementary Figure `r getFigNb()`
Mean clustering F1 score per subpopulation, mean F1 at true number of clusters, as well as maximum and median proportion of excluded cells per subpopulation across various filtering strategies. Doublet removal generally improves clustering accuracy with very mild exclusion rates, even in datasets that do not have heterotypic doublets. Stringent distribution-based filtering creates large cell type biases.

@@

```{r}
res <- readRDS("../data/selection_old.rds")
res <- res[grep("deviance|seurat.variance.standardized",res$sel),]
res$filt <- as.factor(res$filt)
levels(res$filt) <- c("coding only", "no mitochondrial", "no ribosomal", "all")
res2 <- res[which(abs(res$nbClusters-datasets[res$dataset,"subpopulations"])<=1),]
ll <- lapply(split(res2, res2$norm), FUN=function(x){
  m2 <- cast2(x, formula=filt~dataset, fun.aggregate=median, value.var="ARI")
})
chm(ll, scale="column", scaleTitle="z-score", sameScale=TRUE, value_format="%.2f", row_sortFn=function(x,...){ median(x,...) }, cluster_columns=FALSE)
```

### 2020-01-29 supplementary_figures/supp3_normalization.Rmd
updated supplementary figures

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(reshape2)
  library(dplyr)
})
suppressMessages(devtools::load_all("../../pipComp/"))
data(datasets)
source("../misc_functions.R")
```

@@

```{r norm_r2, fig.width=8, fig.height=9}
res <- readRDS("../../resNew/norm2_intermediateSummary.rds")
r2 <- res$dimreduction$PC.R2
r2$norm <- gsub("norm.","",reduceMethodNames(r2$method,3),fixed=T)
r2 <- r2[which(r2$dim <= 10),]
r2$R.on.Clusters <- sqrt(r2$value)
r3 <- cast2(r2, formula=norm~dataset, value.var="R.on.Clusters")
r2m <- dplyr::summarise(group_by(r2,dataset), meanR=mean(R.on.Clusters))
p1 <- ggplot(r2, aes(norm, R.on.Clusters, fill=norm)) + geom_violin() + facet_wrap(~dataset, nrow=1) + coord_flip() + theme(legend.position="none") + stat_summary(fun.y="mean", geom="point") + geom_hline(data=r2m, aes(yintercept=meanR)) + ggtitle("Proportion of the top dimensions explained by real clusters") + ylab("sqrt(R^2) of dimension ~ real subpopulation")

pc1.covar <- lapply( res$dimreduction$PC1.covarR[2:3], FUN=function(x){
  x <- aggregate(x, by=list(reduceMethodNames(row.names(x), 3)), FUN=median)
  row.names(x) <- x[,1]
  x[,1] <- NULL
  x
})
names(pc1.covar) <- c("Correlation with log10-counts", "Correlation with nb features")
p2 <- grid.grabExpr(draw(chm(pc1.covar, sameScale=TRUE, scaleTitle="cor", value_format="%.2f", row_sortFn=median)))
plot_grid(p2, p1, labels=LETTERS[1:2], nrow=2)
```

### Supplementary Figure `r getFigNb()`
**A:** Correlation between library size (left) or the number of detected features (right) and the residuals of a linear regression of the first principal component on the subpopulation.
**B:** Proportion of the variance in the first 10 PCA components that is explained by real subpopulations. The square-root of R-squared values is plotted for ease of visualization.

\newpage

```{r norm_silwidth, fig.width=8, fig.height=7}
res <- readRDS("../../resNew/norm2_intermediateSummary.rds")
sw <- res$dimreduction$clust.avg.silwidth[[1]]
colnames(sw) <- gsub("norm.","",reduceMethodNames(colnames(sw), 3),fixed=T)
sw <- sw[order(reduceMethodNames(row.names(sw),1), rowMeans(sw)),]
ra <- data.frame(meanSilWidth=rowMeans(sw), dataset=reduceMethodNames(row.names(sw),1))
chm(list("Cluster mean silhouette width"=sw), scaleTitle="row z-score", scale="row", row_ann=ra, show_row_names=FALSE, col_sortFn=mean, cluster_rows=FALSE)
```

### Supplementary Figure `r getFigNb()`
Mean silhouette width of true subpopulations across the first 10 components with different normalization and scaling approaches.

\newpage

>>>>

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(reshape2)
  library(dplyr)
  library(viridis)
})
suppressMessages(devtools::load_all("../../pipComp/"))
data(datasets)
theme_set(theme_cowplot(font_size = 11))
source("../misc_functions.R")
```

@@

```{r norm_nbClusters, fig.width=8, fig.height=8.5}
res <- readRDS("../data/norm2_endSummary.rds")
res2 <- res[which(res$filt=="deffilter" & res$norm %in% c("norm.scran", "norm.scran.noscale", "norm.seurat", "norm.seuratvst", "norm.stableGsum")),]
res2$diff.nbClusters <- res2$nbClusters-datasets[res2$dataset,"subpopulations"]
res3 <- aggregate(res2[,"diff.nbClusters"], by=res2[,c("dataset","dims","resolution","norm")], FUN=mean)
colnames(res3)[5] <- "nb"
res3$norm <- gsub("norm.","",res3$norm, fixed=T)
# ggplot(res3, aes(factor(dims), factor(resolution), fill=nb)) + stat_bin2d() + 
#   scale_fill_viridis() + ylab("Resolution") + xlab("Number of dimensions") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#   coord_flip() + facet_grid(dataset~norm)

ll <- lapply(split(res3, res3$dataset), FUN=function(x){
  ggplot(x, aes(factor(resolution), factor(dims), fill=nb)) + stat_bin2d() + 
    scale_fill_viridis() + xlab("Resolution") + ylab("dims") +
    facet_grid(.~norm)
})
for(i in names(ll)){
  ll[[i]] <- ll[[i]] + ggtitle(i)
  if(i==names(ll)[length(ll)]){
    ll[[i]] <- ll[[i]] + theme(axis.text.x = element_text(angle = 90, hjust = 1))
  }else{
    ll[[i]] <- ll[[i]] + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
  }
}
ll[[4]] <- ll[[4]] + ylab("Number of dimensions")

mod <- summary(lm(nbClusters~dataset+norm+dims+resolution, data=res2))
plot_grid(plotlist = ll[-2], nrow=6, rel_heights=c(1,1,1,1,1,1.5))
```

### Supplementary Figure `r getFigNb()`
Mean difference between the number of detected clusters and the number of real subpopulations, depending on the normalization method, the resolution and the number of dimensions used. The Kumar dataset is not shown here due to a lack of variation in the number of clusters detected. A rough ANOVA on `nbClusters~dataset+norm+dims+resolution` suggests that seuratvst (sctransform) is associated with a higher number of clusters (p~`r round(mod$coefficients["normnorm.seuratvst",4],3)`).

### 2020-01-29 supplementary_figures/supp4_selection.Rmd
updated supplementary figures

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}
```

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
})
suppressMessages(devtools::load_all("../../pipComp/"))
data(datasets)
source("../misc_functions.R")
```

@@

layout(matrix(1:4,nrow=2,byrow=TRUE))
LSD::heatscatter(gi$mixology10x3cl$vst.R2, gi$mixology10x3cl$lognorm.R2, xlab="Prop variance explained based, sctransform", ylab="Prop variance explained, standard seurat normalization", main="A. Variance explained by real subpopulations (mixology10x3cl)")

LSD::heatscatter(gi$mixology10x3cl$devianceExplained, gi$mixology10x3cl$vst.R2, xlab="Deviance explained by real subpopulations", ylab="sctransform-based variance explained", , main="B. Variance explained vs deviance explained (mixology10x3cl)")

diffplot <- function(x, title=""){
  LSD::heatscatter(x$log10_mean_counts, x$devianceExplained - x$vst.R2, xlab="log10(mean counts)", ylab="Prop deviance explained - prop variance explained", main=title)
}
diffplot(gi$mixology10x3cl, "C. Difference between the estimates (mixology10x3cl)")
diffplot(gi$Zhengmix4eq, "D. Difference between the estimates (Zhengmix4eq)")
```

@@

```{r, fig.width=8, fig.height=9, warnings=FALSE, messages=FALSE}
ll <- lapply(names(gi), FUN=function(x){
  cumVE2(gi[[x]], gi[[x]]$vst.R2) + xlim(0,2000) + ggtitle(x) + ylab("Cumulative var explained") + theme(legend.position="none")
})
#ggpubr::ggarrange(plotlist=ll, common.legend = TRUE, ncol=2, nrow=4, labels=LETTERS[1:length(ll)])
ll$legend <- get_legend(ll[[1]] + theme(legend.position="right"))
plot_grid(plotlist=ll, ncol=2)
```

@@

```{r, fig.width=8, fig.height=9, warnings=FALSE, messages=FALSE}
ll <- lapply(names(gi), FUN=function(x){
  cumVE2(gi[[x]], gi[[x]]$devianceExplained) + xlim(0,2000) + ggtitle(x) + ylab("cumul deviance explained") + theme(legend.position="none")
})
ll$legend <- get_legend(ll[[1]] + theme(legend.position="right"))
plot_grid(plotlist=ll, ncol=2)
```

@@

```{r, fig.width=8, fig.height=6}
res <- readRDS("../../resNew/sel_intermediateSummary.rds")
r2 <- res$dimreduction$PC.R2
r2 <- r2[grep("1000$",r2$method),]
r2 <- r2[grep("doublet.pl deffilter",r2$method,fixed=T),]
r2$norm <- reduceMethodNames(r2$method,3)
r2$method <- reduceMethodNames(r2$method,4)
r2 <- r2[which(r2$dim <= 5),]
r2$R.on.Clusters <- sqrt(r2$value)
r3 <- lapply(split(r2, r2$norm), FUN=function(x){ 
  cast2(x, formula=method~dataset, value.var="R.on.Clusters")
})
chm(r3[-1], scale="column", scaleTitle="z-score", value_format="%.2f", row_sortFn=median, sameScale=TRUE)
```

@@

Proportion of the variance in the first 5 principal components that is explained by real subpopulations according to different normalization and selection procedures, selecting 1000 genes in each case.

>>>>

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}

suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
})
suppressMessages(devtools::load_all("../../pipComp/"))
data(datasets)
source("../misc_functions.R")
```

@@

layout(matrix(1:4,nrow=2,byrow=TRUE))
par(mar= c(7, 4, 5, 2) + 0.1)
LSD::heatscatter(gi$mixology10x3cl$vst.R2, gi$mixology10x3cl$lognorm.R2, xlab="Prop var explained based, sctransform", ylab="Prop var explained, lognorm", main="A. Var explained by subpopulations\n(mixology10x3cl)", cex.main=0.8)

LSD::heatscatter(gi$mixology10x3cl$devianceExplained, gi$mixology10x3cl$vst.R2, xlab="Deviance explained by subpopulations", ylab="sctransform-based variance explained", , main="B. Var explained vs deviance explained\n(mixology10x3cl)", cex.main=0.8)

diffplot <- function(x, title=""){
  LSD::heatscatter(x$log10_mean_counts, x$devianceExplained - x$vst.R2, xlab="log10(mean counts)", ylab="Prop deviance explained - prop variance explained", main=title, cex.main=0.8)
}
diffplot(gi$mixology10x3cl, "C. Difference between the estimates\n(mixology10x3cl)")
diffplot(gi$Zhengmix4eq, "D. Difference between the estimates\n(Zhengmix4eq)")
```

@@

```{r, include=FALSE}
ll <- lapply(names(gi), FUN=function(x){
  cumVE2(gi[[x]], gi[[x]]$vst.R2) + xlim(0,2000) + ggtitle(x) + ylab("Cumulative var explained") + theme(legend.position="none")
})
#ggpubr::ggarrange(plotlist=ll, common.legend = TRUE, ncol=2, nrow=4, labels=LETTERS[1:length(ll)])
ll$legend <- get_legend(ll[[1]] + theme(legend.position="right"))
```

```{r, fig.width=8, fig.height=9, results="hide", warnings=FALSE, messages=FALSE}
plot_grid(plotlist=ll, ncol=2)
```

@@

```{r, include=FALSE}
ll <- lapply(names(gi), FUN=function(x){
  cumVE2(gi[[x]], gi[[x]]$devianceExplained) + xlim(0,2000) + ggtitle(x) + ylab("cumul deviance explained") + theme(legend.position="none")
})
ll$legend <- get_legend(ll[[1]] + theme(legend.position="right"))
plot_grid(plotlist=ll, ncol=2)
```

```{r, fig.width=8, fig.height=9, results="hide", warnings=FALSE, messages=FALSE}
plot_grid(plotlist=ll, ncol=2)
```

@@

```{r, fig.width=8, fig.height=9}
res <- readRDS("../data/results_selection.rds")
cl <- res$clustering
cl <- cl[grep("devianceExplained|varExp", row.names(cl), invert=TRUE),]
vst <- cl[grep("seuratvst",row.names(cl)),]
snorm <- cl[grep("seuratvst",row.names(cl), invert=TRUE),]

ds <- sapply(strsplit(colnames(cl)," "), FUN=function(x) x[1])
ds <- split(seq_len(ncol(cl)), ds)
ds <- ds[grep("stepElapsed",names(ds),invert=TRUE)]
myag <- function(cl, out="ARI", trueK=FALSE){
  y <- lapply(ds, FUN=function(x){
    x <- cl[,x]
    colnames(x) <- sapply(strsplit(colnames(x)," "), FUN=function(x) x[2])
    x <- parsePipNames(x)
    if(!trueK) return( aggregate( x[,out], 
                                  by=list(method=x$sel, nb=x$selnb), 
                                  na.rm=TRUE, FUN=mean) )
    w <- which(x$n_clus==x$true.nbClusts)
    aggregate(x[w,out], by=list(method=x$sel[w], nb=x$selnb[w]), na.rm=TRUE, FUN=mean)
  })
  bind_rows(y, .id="dataset")
}
vst.ARIatK <- myag(vst, trueK=TRUE)
snorm.ARIatK <- myag(snorm, trueK=TRUE)

cols <- getQualitativePalette(9)
names(cols) <- unique(vst.ARIatK$method)

plot_grid( 
  ggplot(vst.ARIatK, aes(nb, x, colour=method)) + geom_line(size=1.3) +facet_wrap(~dataset, scales="free") + scale_color_manual(values = cols) + xlab("Number of genes selected") + ylab("ARI at true # of clusters") + theme(axis.text=element_text(size=10)),
  ggplot(snorm.ARIatK, aes(nb, x, colour=method)) + geom_line(size=1.3) +facet_wrap(~dataset, scales="free") + scale_color_manual(values = cols) + xlab("Number of genes selected") + ylab("ARI at true # of clusters") + theme(axis.text=element_text(size=10)),
  nrow=2, labels = LETTERS[1:2]
)
```

@@

Clustering accuracy according to the number of genes selected using various ranking/selection methods. **A:** Based on sctransform, **B:** Based on standard Seurat normalization.

### 2020-01-29 supplementary_figures/supp5_dr.Rmd
updated supplementary figures

```{r, fig.width=8, fig.height=8}
res <- readRDS("../data/dr_intermediateSummary.rds")
sw <- res$dimreduction$clust.avg.silwidth[[1]]
colnames(sw) <- reduceMethodNames(colnames(sw),c(3,4,6))
colnames(sw) <- gsub("norm.","",colnames(sw),fixed=T)
sw <- matrix(unlist(sw),ncol=ncol(sw), dimnames=dimnames(sw))
df <- data.frame(Norm=reduceMethodNames(colnames(sw),1), Selection=reduceMethodNames(colnames(sw),2), DimRed=reduceMethodNames(colnames(sw),3), stringsAsFactors=FALSE)
df$Norm[grep("GlmPCA", df$DimRed)] <- ""
ag <- aggregate(t(sw), by=df, FUN=median)
ag <- ag[order(ag[,3],ag[,1],ag[,2]),]
sw2 <- t(ag[,-1:-3])
colnames(sw2) <- apply(ag[,c(1,3)],1,collapse=" ",FUN=paste)

# hma <- HeatmapAnnotation(df=data.frame(meanSilWidth=rowMeans(sw2), dataset=reduceMethodNames(row.names(sw),1)))
# chm(list("Cluster mean silhouette width"=t(sw2)), scaleTitle="meanSilWidth", scale="none", show_row_names=TRUE, show_column_names=FALSE, row_ann=ag[,1:3], top_annotation=hma, cluster_rows=FALSE)

hma <- HeatmapAnnotation(df=ag[,1:3])
chm(list("Cluster mean silhouette width"=sw2), scaleTitle="meanSilWidth", scale="none", row_ann=data.frame(meanSilWidth=rowMeans(sw), dataset=reduceMethodNames(row.names(sw),1)), show_row_names=FALSE, top_annotation=hma)
```

### Supplementary Figure `r getFigNb()`
Average silhouette width of real subpopulations when using the first 10 components of different dimensionality reductions. Weighing the components according to their variance increases the distinctiveness of the subpopulations. Nevertheless, Seurat's standard PCA appears superior to the alternatives.

```{r, fig.width=8.5, fig.height=5}
r2 <- res$dimreduction$PC.R2
r2$norm <- gsub("norm.","",reduceMethodNames(r2$method,6),fixed=T)
r2 <- r2[which(r2$dim <= 10),]
r2$R.on.Clusters <- sqrt(r2$value)
r3 <- cast2(r2, formula=norm~dataset, value.var="R.on.Clusters")
r2m <- dplyr::summarise(group_by(r2,dataset), meanR=mean(R.on.Clusters))
ggplot(r2, aes(norm, R.on.Clusters, fill=norm)) + geom_violin() + facet_wrap(~dataset, nrow=1) + coord_flip() + theme(legend.position="none") + stat_summary(fun.y="mean", geom="point") + geom_hline(data=r2m, aes(yintercept=meanR)) + ggtitle("Proportion of the top dimensions explained by subpopulations") + ylab("sqrt(R^2) of dimension ~ subpopulation") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Supplementary Figure `r getFigNb()`

Proportion of the variance in the first 10 principal components that is explained by real subpopulations according to different dimensionality reduction methods.

```{r}
res <- readRDS("../data/dr_endSummary.rds")
res2 <- res[which(abs(res$nbClusters-datasets[res$dataset,"subpopulations"])<=1),]
ll <- lapply(split(res2, res2$norm), FUN=function(x){
  m2 <- cast2(x, formula=dr~dataset, fun.aggregate=median, value.var="ARI")
})

chm(ll, scale="column", scaleTitle="z-score", sameScale=TRUE, value_format="%.2f", row_sortFn=function(x){ median(x,na.rm=T) }, cluster_columns=FALSE)
```

### Supplementary Figure `r getFigNb()`
ARI of the clustering based on different normalization and dimensionality reduction procedures.

@@

```{r, fig.width=8, fig.height=8.5}
library(viridis)
res <- readRDS("../data/clust_endSummary.rds")
res2 <- res[which(res$clust=="clust.seurat" & res$k==20 & res$dims>2 & res$resolution > 0.05 & res$resolution <= 1),]
res2 <- res2[which(res2$dr=="seurat.pca" & res2$dataset!="Kumar"),]
res3 <- aggregate(res2[,"ARI"], by=res2[,c("dataset","dims","resolution")], FUN=mean)
ll <- lapply(split(res3, res3$dataset), FUN=function(x){
  ggplot(x, aes(factor(dims), factor(resolution), fill=ARI)) + stat_bin2d() + 
    scale_fill_viridis() + ylab("Resolution") + xlab("Number of dimensions") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
})
for(i in names(ll)) ll[[i]] <- ll[[i]] + ggtitle(i)
plot_grid(plotlist = ll, nrow=3)
```

>>>>

@@

```{r, fig.width=8, fig.height=8.5}
library(viridis)
res <- readRDS("../data/clust_endSummary.rds")
res2 <- res[which(res$clust=="clust.seurat" & res$k==20 & res$dims>2 & res$resolution > 0.05 & res$resolution <= 1),]
res2 <- res2[which(res2$dr=="seurat.pca" & res2$dataset!="Kumar"),]
res3 <- aggregate(res2[,"ARI",drop=FALSE], by=res2[,c("dataset","dims","resolution")], FUN=mean)
ll <- lapply(split(res3, res3$dataset), FUN=function(x){
  ggplot(x, aes(factor(dims), factor(resolution), fill=ARI)) + stat_bin2d() + 
    scale_fill_viridis() + ylab("Resolution") + xlab("Number of dimensions") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
})
for(i in names(ll)) ll[[i]] <- ll[[i]] + ggtitle(i)
plot_grid(plotlist = ll, nrow=3)
```
